Synthesis of Statically Analyzable Accelerator Networks From Sequential Programs

### 2016-04-24 paper.tex
Updates from ShareLaTeX

\numberofauthors{0} %  in this sample file, there are a *total*
% of EIGHT authors. SIX appear on the 'first-page' (for formatting
% reasons) and the remaining two appear in the \additionalauthors section.
%
\begin{comment}
\author{
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
%
% 1st. author
\alignauthor
Shaoyi Cheng \\%\titlenote{Dr.~Trovato insisted his name be first.}\\
       \affaddr{EECS, University of California, Berkeley}\\
       %\affaddr{1932 Wallamaloo Lane}\\
       %\affaddr{Wallamaloo, New Zealand}\\
       \email{sh\_cheng@berkeley.edu}
% 2nd. author
\alignauthor
John Wawrzynek \\%\titlenote{The secretary disavows any knowledge of this author's actions.}\\ 
       \affaddr{EECS, University of California, Berkeley}\\
       %\affaddr{P.O. Box 1212}\\
       %\affaddr{Dublin, Ohio 43017-6221}\\
       \email{johnw@eecs.berkeley.edu}
% 3rd. author
}
\end{comment}
\maketitle
\begin{abstract}
This paper describes a general framework in
transforming a sequential program into a network of processes, which are
converted to hardware accelerators through high level synthesis.
A complementing technique is also proposed to statically analyze the behaviors of the 
generated accelerator network.
The interactions between the accelerators' schedules, the capacity of communication
channels in the network and the memory access mechanisms are incorporated
into our model, such that potential artificial deadlocks can be detected and resolved \textit{a priori}.

@@

\keywords{high level synthesis, process network, system design flow,  parallelism, hardware accelerator}

\section{Introduction}
As the growth of processor clock frequency came to a stop in the middle of the last decade, the performance boost provided by the continuous advancement of hardware technology for sequential programs could no longer be sustained. The benefits of user transparent CPU architecture optimizations were also diminishing. With the recent advent of FPGA SoCs~\cite{chips:zynq}, offloading computation to hardware accelerators becomes a promising alternative for
achieving better system efficiency. However,
%To create good implementation in the spatial computing
%paradigm however, is far less intuitive compare to imperative programming.
there is a big gap between the inherently parallel mode of operation in hardware accelerators and the intuitive imperative programming paradigm. Even with tool flows like
high level synthesis (HLS), substantial user inputs are still required to create
efficient implementations~\cite{7082747}. To address this problem, we envision a two
step process, where the sequential specification provided by the user is first
transformed to a more parallel representation, from which the conventional HLS
can easily create high performance compute engine.

@@

%In this paper, we explore the transformation from sequential programs to networks of processes.
In this paper, we explore this approach by using a modified version of process network (PN), a parallel model of computation, as the target representation for our transformation.
Using FPGA as the computing device, our final implementations are accelerator networks optimized to achieve high throughput in the presence of long data access latency.
Our major contributions include:
\begin{itemize}
    \item We devise a general approach for converting a sequential program to a network
of concurrent processes. Different algorithms can potentially be plugged into the flow, splitting computations between processes in different ways, optimized for the underlying computing substrates.
    \item With FPGA as our target platform, processes are to be converted to hardware accelerators. We
    propose an analysis framework to statically detect and resolve deadlocks in the accelerator networks, ensuring the correctness of the execution. 
    \item We develop an algorithm to generate instances of process networks
particularly amenable to hardware acceleration. Using high level synthesis (HLS), we mapped them onto FPGAs
to demonstrate the benefits of our approach. 
\end{itemize}

@@

The rest of the paper is organized as follows:
Section~\ref{pre} provides preliminaries in process network
%intermediate representation of the sequential program,
and accelerator generation using HLS. Section~\ref{mainConvert} describes the main steps in transforming sequential programs
to networks of processes. When implemented on FPGAs, these networks may experience deadlock issues, which can be analyzed and resolved with the method developed in section~\ref{deadlock}. To make our general approach more concrete, we presents a specific algorithm which fits into our PN generation framework in section~\ref{exampleAlgo}.
This algorithm and its complementing optimizations are developed with FPGA as the target platform, thus the generated networks boast great performance when mapped onto the intended 
computing substrate (section~\ref{expEval}). Finally, related work is presented in section~\ref{relatedwork}, before we conclude the paper (section~\ref{conclude}).
%with discussion and future directions (section~\ref{conclude}).

\section{Preliminaries}
\label{pre}
\begin{comment}
A typical program written for processor execution 
is sequential in its semantics. Techniques to
parallelize this specification need to preserve
the resultant behaviors of the original ordering
of the computation and data access operations.
In this work, we limit the scope of our study
to functions implemented in imperative
languages who terminate in finite time, so that
empirical comparisons can be made. Our approach
tries to transform functions to a different 
model of computation, process networks, before
their
actual implementation on physical computing devices.
\end{comment}

\subsection{Process Networks}
\label{pre1}
The process network, or Kahn process network~\cite{gilles1974semantics}, is an inherently
parallel model of computation where processes communicate with
each other through unbounded FIFO channels. 
The processes cannot test for availability of data tokens in 
the channels without consuming them. If a channel is empty
the process blocks as it reads from the channel.
Writing to the FIFOs, however, is non-blocking.
In addition, each FIFO can only be consumed by a single process, and be written to by a single process. The PN model is deterministic in the sense that the scheduling of
the process execution does not alter the final results. This gives great flexibility
in implementing/controlling each individual processes.

A characteristic of PN is that all data exchanges are performed over the FIFOs and the notion of a shared global memory does not exist. A program executing on 
a processor however, always have instructions communicating with each other temporally using
the memory. In this work, we bridge this difference by modeling data access as inter-process communication.
%such that our deadlock analysis can also take into account the accelerators' interaction with the memory. 
More specifically, the memory, and its associated interconnect can be seen as a special process which takes in address/data tokens and produces data/acknowledge tokens. In the case where multiple memory ports are exposed, this ``process" does not block on read, as it can choose to serve any port with valid
requests. This behavior is similar to the ``select" function in YAPI~\cite{Kock00yapi:application}, which introduced non-determinism. In our framework however, the non-determinism is addressed by a set of special mechanisms, which will be explained in section~\ref{mainConvert:mem}.

Another necessary deviation from the ideal PN model lies in the capacity of the channels.
In any practical implementations of process network, the FIFOs cannot be unbounded. 
Consequently, writes may block and deadlocks can potentially happen. 
These are so-called artificial deadlocks, as their occurrence is introduced by the nonideality of the implementation and the circular dependency formed %in these deadlocks 
always contain operations writing to a full FIFO.
%Note the memory ``process", which does not block on reads,
%can block on writes (responses to other processes), when a large amount of data returned are not consumed promptly. It may therefore also contribute to the development of artificial deadlocks. 
In a process network, % generated from a sequential program, 
these deadlocks, if left unresolved, would cause a mismatch between
the resultant behavior of the original code and the network. This issue is addressed by our
analysis technique described in section~\ref{deadlock}.

@@

\subsection{High Level Synthesis of Accelerators}
\label{pre2}
A typical HLS tool attempts to capture parallelism in a function written with high level languages like C/C++. With the help of various user provided guidance, it generates hardware in the form of RTL.
Compute operations and data accesses are allocated and scheduled according to the dependency constraints extracted from the input program, and the resource constraints imposed by the target platform. The circuits generated usually follow a Finite State Machine with Datapath (FSMD) paradigm. Activation of a particular operator in the datapath is associated
with a certain clock cycle and the execution of the computation engine is orchestrated by a synthesized central controller.

@@

The insertion of the sender/receiver of the barrier happens after the instruction partitioning. Depending on if a partition is the sender or receiver of the token,
an extra store/load operation is introduced into the set of instructions, associated
with the basic block where the barrier occurs.
\end{comment}

\subsubsection{Optimizing Communication v.s. Computation}
\label{opt}
 Regardless of exactly how the instructions are partitioned, as each process performs its computation locally, the necessity to communicate with its peers can incur implementation overheads.
For each communication channel, buffer space is to be allocated, and sender/receiver primitives also need to be inserted. Therefore, it is sometimes beneficial to duplicate some nodes after the instruction partitioning. Given sets of instruction nodes created after applying a platform-specific algorithm, for each of these set $G$, the problem of whether to copy other instruction nodes into $G$ can be formulated as an integer programming problem.

@@

Figure~\ref{fig:example} is a simple generated PN we use to illustrate this problem. The original program is converted to a two process network
with two single slot communication channels between them.
Figure~\ref{fig:2schedules} then shows the scheduling and communication of the processes. Each instruction invocation is prefixed with a vector, representing its place in the (j,i) iteration space, and three types of dependency edges are used
to show the necessary precedence among these invocations. There is no cycle of
dependency in (a), and therefore no deadlocks, when both processes execute in program order. However, in (b), 
a common HLS technique, loop pipelining, is applied to \textit{Process 1}, where
iterations of the loop are aggressively overlapped. Every cycle, a new iteration is started. Because of the long latency of load from \textit{in[i]} and
the multiplication, the sending of the first \textit{r2} token, $_{(0,0)}$store(r2Q,r2), is scheduled
after many instruction invocations for later iterations. Due to the static scheduling, it can not execute until after $_{(0,2)}$store(r1Q,r1), whose
writing to FIFO r1Q cannot happen until the previous token is consumed by 
$_{(0,1)}$load(r1Q). Because of the scheduling in process 2, this load from r1Q
occurs after $_{(0,0)}$load(r2Q), which cannot complete unless $_{(0,0)}$store(r2Q,r2)
can send the token.  An artificial deadlock is therefore created.

@@

\begin{figure}[htp]
\begin{center}
\includegraphics[width=0.8\linewidth]{fig/predGraphNew.pdf}
\caption{Precedence Graph for Process Network
\label{fig:predgraph}}
\end{center}
\vspace{-2.0em}
\end{figure}

@@

Resolving
the deadlock involves choosing one of the $DepF$ edges
in the cycle and add enough slots (making the weight more negative)
such that the weight of the cycle gets a negative leading element. For our example in figure~\ref{fig:predgraph}, if the capacity of r1Q
is increased by 1, the weight on its corresponding $DepF$ becomes (0,-2). Consequently, the identified cycle in the figure would have
a total weight (0,-1), when the original artificial deadlock disappears.
In general, for every deadlock causing cycle, the $DepF$ edge whose channel has the smallest width can be selected for capacity expansion, so the cost of these extra slots is minimized. This heuristic does not necessarily produce an absolute
global optimal solution in the presence of multiple deadlock causing cycles, but suffices for our use cases.
\begin{comment}
Of course, as multiple non-negatively weighted cycles can be present simultaneously, the absolute global optimal solution would involve finding a set of $DepF$ edges, and minimize the aggregate cost of incrementing each for all cycles to be resolved. This is an NP hard problem
\end{comment}
%\subsection{Modelling Burst Mode Memory Access}

\begin{comment}

In~\cite{}, KPNs are 
categorized into strictly bounded, bounded and unbounded ones.
A KPN is strictly bounded if and only if any execution of of the network requires bounded space. It is bounded if and only if there are some
execution requiring bounded space while it is unbounded if and only if any
execution requires unbounded space. Methods to find bounds for communication channels
in KPNs were discussed in~\cite{}\cite{}. In general, whether a
KPN is bounded is undecidable~\cite{}, but for our process networks, which
are generated from sequential programs expressed in single static assignment
form, we can analyze the boundedness of the FIFO channels in various
scenarios.

To manifest the same behavior as the original sequential program, the generated network
needs to be bounded. This can be easily proven as a execution schedule constructed according to the original

\end{comment}

\subsection{Deadlocks in Memory Accesses}
To exploit the memory bandwidth more efficiently, burst mode memory accesses are often used as well. Even though the memory ``process" does not block on reads, if the response data it sends to other normal processes is not consumed promptly, it can blocked on writes. 
The head-of-line blocking can then create a dependency cycle, resulting in a deadlock. Our precedence graph can be used to compute necessary buffer size to resolve this situation as well.

Figure~\ref{fig:memBurst} illustrates how multiple burst mode memory accesses can
create a deadlock. The reaction of the memory is modelled by a pop action
responding to the incoming request, and a push action producing one response data token. Meanwhile, the sending of memory requests and receiving of responses are decoupled and assigned to different slots in the process schedule. This is often done in HLS because of the long latency of memory accesses.
For burst mode requests associated with a loop, the request sending is moved to outside of the loop as shown in the figure. The symbolic variable $B$ in the weight of the $DepM$ edge signifies how many tokens have to be consumed/buffered before the response for the next request ($data2$) become accessible for the response receiver. Assume the arbiter for the memory subsystem implements a policy which can potentially take in and serve multiple ($K$) consecutive requests from one stream, $B$ can be as big as $BurstSize_{max}*K$. 
In the case of a round-robin arbiter, $B = BurstSize_{max}$ and for our experiment platform which uses AXI bus protocol, this number is 256. With the precedence graph, we can easily compute that the buffering for the two memory response channels needs to be increased to $B$ for the deadlocks to be resolved.
\begin{figure}[htp]
\begin{center}
\includegraphics[width=1.05\linewidth]{fig/burstMem.pdf}
\caption{Deadlock in Burst Memory Accesses
\label{fig:memBurst}}
\end{center}
\vspace{-2.0em}
\end{figure}

@@

\begin{comment}
The benchmarks data footprints of the benchmarks
are listed in table~\ref{tab:datasize}

To evaluate the benefits of our approach, we took several benchmark kernels
and processed them with our flow. For sparse matrix vector (SpMV) multiply---our
first kernel---compressed sparse
row (CSR) format is used to store the matrix, where the loads of the floating
point numbers to be multiplied depend on the data in an index array.
The next two kernels, Knapsack and Floyd-Warshall, are both dynamic programming problems. The memory
addresses to be accessed are derived from the results of computation.
Our final kernel, Depth first search (DFS), is a widely used graph algorithm. The benchmarked version
operates on pointer based data structure and uses a stack. All these kernels
are sequential code with non-regular control flow and memory access patterns. 
The input dataset for each benchmark, as described by 
Table~\ref{tab:datasize}, is also too big to fit entirely on chip. 
Due to the presence of statically unknown off-chip memory accesses, these benchmarks are good cases where our algorithm can play a significant role in improving the overall performance.
On the other hand, for problems where the entire data set can be buffered on chip or when the memory access patterns are regular, our approach would offer little
performance advantage over the conventional DMA+accelerator approach~\cite{vivado_hls:appnoteMMult}.
\end{comment}

@@

The FPGA device used for our evaluation is Xilinx's Zynq-7000 XC7Z020 FPGA
SoC. It contains two parts: an ARM based processing
system (PS), and the programmable logic (PL). 
The baseline for our evaluation is the performance of each kernel executing
on the ARM core, which is an out-of-order, dual-issue hard processor running at 667MHz. Zynq also provides two
options for the accelerators in PL to access the main memory: 
through the accelerator coherence port (ACP), or the high performance (HP) port. 
ACP connects to the snoop control unit in the PS, and 
thus accesses the on-chip cache of the PS. On the other hand, the HP port connects
directly to the memory controller, thus often incurs longer latency
for data requests.
\begin{comment}
which necessitates the flushing of cache lines 
by the processor if cached memory is accessed by the accelerator.
In either case, if memory states are also buffered in the PL with caches, they 
need to be explicitly pushed to the processing system side after the accelerator finishes running. 
As both the ACP
and the HP are slave ports, they provide no mechanisms to extract data from the programmable logic
when the ARM processor is running. The interaction between the generated accelerators
and the main pieces of the FPGA SoC is shown in figure~\ref{fig:ippf}.
\end{comment}

@@

\end{tabular}
\label{tab:datasize}
\end{table}

\begin{comment}
\begin{figure}[htp]
\begin{center}
\includegraphics[width=0.9\linewidth]{fig/system.pdf}
\caption{Implementation of Dataflow Accelerators in FPGA SoC  
\label{fig:ippf}}
\end{center}
%\vspace{-1.0em}
\end{figure} 
\end{comment}

\begin{figure*}[htp]
\begin{center}
\includegraphics[width=0.95\linewidth]{fig/perfNew.pdf}
\caption{Performance of Conventional Accelerators and Accelerator Networks, normalized to the hard ARM core. Implementations with caches use
Xilinx System Cache IP, configured to be 64KB and 2-way associative   
\label{fig:acpraw}}
\end{center}
\vspace{-2.7em}
\end{figure*}

In our study, Vivado HLS, a state-of-the-art high level synthesis tool provided by Xilinx,
is used to create the accelerators for the original sequential code, as well as the processes generated by our flow. With the target clock period set to 
8ns during HLS, the tightest timing constraints post place \& route implementations managed to meet range from 111 to 150MHz. 
All design points shown in this section 
use the highest achievable frequency as the actual operating clock frequency. 
%\vspace{-1.0em}
\subsection{Performance Comparisons}
In figure~\ref{fig:acpraw}, 
performance of the different implementations are presented. 
Conventional accelerators and accelerator networks with different memory subsystem
configurations are compared. All the numbers are normalized to the baseline.

@@

\begin{comment}
It is also apparent that our approach has its limitations, as demonstrated by its ineffectiveness in the 
depth first search benchmark.
The kernel performs very little computing but lots of memory accesses.
The use of a stack in DFS also creates a dependence cycle through the memory
and consequently, the performance is fundamentally limited by the latency of memory access.
Thus there were only small differences between the performance of the conventional accelerator and the dataflow accelerator. Besides, the memory access pattern
does not provide many opportunities for optimizations. As a result, both accelerators achieves performance far below that of the baseline.
\end{comment}

@@

%\subsection{Area and Energy Comparison}
%\vspace{-1.0em}
\subsection{Area comparison}
To quantify the impact of our proposed methodology on area, 
we have compared the FPGA resouce usage of conventional accelerators and
the accelerator networks.
Table~\ref{tab:areacom} shows the results, 
where each acclerator is complemented with two different memory subsystem configurations.
\renewcommand{\tabcolsep}{0.5pt}
\begin{table}[htbp]
%\vspace{-1.2em}
\caption{Resource Usage of Accelerators.}
%\vspace{-1.2em}
\scriptsize
\centering
\begin{tabular}{| c | c | c | c | c | c | c | c| }
  \hline            
  \multirow{2}{*}{} &  &   \multicolumn{3}{c|}{\bf ACP  } & \multicolumn{3}{c|}{\bf ACP + 64KB Cache }   \\
 \cline{3-8} 
 {\bf Benchmark}   &    & LUT& FFs& BRAM & LUT &   FFs      & BRAM    \\
  %
  %                 & LUT &FFs&  BRAM &  LUT &   FFs      & BRAM   \\
  \hline            
  \hline

>>>>

\numberofauthors{0} %  in this sample file, there are a *total*
% of EIGHT authors. SIX appear on the 'first-page' (for formatting
% reasons) and the remaining two appear in the \additionalauthors section.
%
\begin{comment}
\author{
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
%
% 1st. author
\alignauthor
Shaoyi Cheng \\%\titlenote{Dr.~Trovato insisted his name be first.}\\
       \affaddr{EECS, University of California, Berkeley}\\
       %\affaddr{1932 Wallamaloo Lane}\\
       %\affaddr{Wallamaloo, New Zealand}\\
       \email{sh\_cheng@berkeley.edu}
% 2nd. author
\alignauthor
John Wawrzynek \\%\titlenote{The secretary disavows any knowledge of this author's actions.}\\ 
       \affaddr{EECS, University of California, Berkeley}\\
       %\affaddr{P.O. Box 1212}\\
       %\affaddr{Dublin, Ohio 43017-6221}\\
       \email{johnw@eecs.berkeley.edu}
% 3rd. author
}
\end{comment}
\maketitle
\begin{abstract}
This paper describes a general framework in
transforming a sequential program into a network of processes, which are
converted to hardware accelerators through high level synthesis.
A complementing technique is also proposed to perform static deadlock analysis for the generated accelerator network.
The interactions between the accelerators' schedules, the capacity of communication
channels in the network and the memory access mechanisms are all incorporated
into our model, such that potential artificial deadlocks can be detected and resolved \textit{a priori}.

@@

\keywords{process network, high level synthesis, system design flow, hardware accelerator}

\section{Introduction}
As the growth of processor clock frequency came to a stop in the middle of the last decade, the performance boost provided by the continuous advancement of hardware technology for sequential programs could no longer be sustained. The benefits of user transparent CPU architecture optimizations were also diminishing. With the recent advent of FPGA SoCs~\cite{chips:zynq}, offloading computation to hardware accelerators becomes a promising alternative for
achieving better system efficiency. However,
%To create good implementation in the spatial computing
%paradigm however, is far less intuitive compare to imperative programming.
there is a big gap between the inherently parallel mode of operation in hardware accelerators and the intuitive imperative programming paradigm. Even with tool flows like
high level synthesis (HLS), substantial user inputs are still required to create
efficient implementations~\cite{7082747}. To address this problem, we envision a two
step process, where the sequential specification provided by the user is first
transformed to a more parallel representation, from which the conventional HLS
can easily create a high performance compute engine.

@@

%In this paper, we explore the transformation from sequential programs to networks of processes.
In this paper, we explore this approach by using a modified version of process network (PN), a parallel model of computation, as the target representation for our transformation.
Using FPGA as the computing device, our final implementations are accelerator networks optimized to achieve high throughput in the presence of long data access latency.
Our major contributions include:
\begin{itemize}
    \item We devise a general approach for converting a sequential program to a network
of concurrent processes. Different algorithms can potentially be plugged into the flow, splitting computations between processes in different ways, optimized for the underlying compute substrates.
    \item With FPGA as our target platform, processes are to be converted to hardware accelerators. We
    propose an analysis framework to statically detect and resolve deadlocks in the accelerator networks, ensuring the correctness of the execution. 
    \item We develop an algorithm to generate instances of process networks
particularly amenable to hardware acceleration. Using high level synthesis (HLS), we mapped them onto FPGAs
to demonstrate the benefits of our approach. 
\end{itemize}

@@

The rest of the paper is organized as follows:
Section~\ref{pre} looks at background and related work in process network
%intermediate representation of the sequential program,
and accelerator generation using HLS. Section~\ref{mainConvert} describes the main steps in transforming sequential programs
to networks of processes. When implemented on FPGAs, these networks may experience deadlock issues, which can be statically analyzed and resolved with the method developed in section~\ref{deadlock}. To make our general approach more concrete, we presents a specific algorithm which fits into our PN generation framework in section~\ref{exampleAlgo}.
This algorithm is developed with FPGA as the target platform, thus the generated networks boast great performance when mapped onto the intended 
compute substrate (section~\ref{expEval}). Finally, we conclude the paper in section~\ref{conclude}.
%with discussion and future directions (section~\ref{conclude}).

\section{Background and Related Work}
\label{pre}

\begin{comment}
A typical program written for processor execution 
is sequential in its semantics. Techniques to
parallelize this specification need to preserve
the resultant behaviors of the original ordering
of the computation and data access operations.
In this work, we limit the scope of our study
to functions implemented in imperative
languages who terminate in finite time, so that
empirical comparisons can be made. Our approach
tries to transform functions to a different 
model of computation, process networks, before
their
actual implementation on physical computing devices.
\end{comment}

\subsection{Process Networks}
\label{pre1}

\begin{comment}
Parallelizing sequential programs by transforming them to process networks was attempted in multiple past projects. The Compaan flow~\cite{mat2pn} takes Matlab code and generates
process networks. ESPAM~\cite{4378214}\cite{c2stream}, on the otherhand, uses sequential C as the input language and also provides an efficient method to compute the needed FIFO size. They also have extension for implementing the PN on FPGAs. However, the targets of their flow are highly regular kernels with statically analyzable memory access patterns, which are quite different than the benchmarks we have applied our framework to.

Finally, HLS for FPGAs are being active developed in both industry~\cite{tools:vivadohls} and academia~\cite{tools:legup}. Researchers
have also tried different programming paradigms~\cite{Lebedev:2010:MMA:1935936.1936078}
to expose greater amount of parallelism to HLS. Our work complements
HLS by providing a higher level transformation infrastructure which, when
used appropriately, can significantly improve the quality of the generated compute engine.
\end{comment}

The process network, or Kahn process network~\cite{gilles1974semantics}, is an inherently parallel model of computation where processes communicate with
each other through unbounded FIFO channels.
The processes block when they read from empty channels but never block when writing.
%The processes cannot test for availability of data tokens in 
%the channels without consuming them. If a channel is empty
%the process blocks as it reads from the channel.
%Writing to the FIFOs, however, is non-blocking.
%In addition, each FIFO can only be consumed by a single process, and be written to by a single process.
The PN model is deterministic in the sense that the scheduling of
the process execution does not alter the final results. This gives great flexibility
in implementing/controlling each individual processes.

%A characteristic of PN is that all data exchanges are performed over the FIFOs and the notion of a shared global memory does not exist. A program executing on 
%a processor however, always have instructions communicating with each other temporally using
%the memory. In this work, we bridge this difference by modeling data access as inter-process communication.

A characteristic
of KPN is that all data exchange are performed over the FIFOs and the notion of
a shared global memory does not exist in the model. Previous works
converting sequential programs to process networks~\cite{mat2pn}\cite{c2stream}
were able to convert original memory operations to explicit data communication
at the boundary of the generated hardware as the targeted kernels are highly 
regular. In this study however, we target kernels with irregular memory accesses and the final implementations are to be used along side a general purpose processor. We therefore preserve the shared memory, but conceptually model data accesses as inter-process communications.
%the memory model needs to be preserved. It is therefore necessary to conceptually
%translate the behavior of processes accessing memory to that of inter-process
%communication.
%such that our deadlock analysis can also take into account the accelerators' interaction with the memory. 
More specifically, the memory, and its associated interconnect can be seen as a special process which takes in address/data tokens and produces data/acknowledge tokens. In the case where multiple memory ports are exposed, this ``process" does not block on read, as it can choose to serve any port with valid
requests. This behavior is similar to the ``select" function in YAPI~\cite{Kock00yapi:application}, which introduced non-determinism. In our framework however, the non-determinism is addressed by a set of special mechanisms, which will be explained in section~\ref{mainConvert:mem}.

Another necessary deviation from the ideal PN model lies in the capacity of the channels.
In any practical implementations of process network, the FIFOs cannot be unbounded.
Consequently, writes may block and deadlocks can potentially happen. 
These are so-called artificial deadlocks, as their occurrence is introduced by the nonideality of the implementation and the circular dependency formed %in these deadlocks 
always contain operations writing to a full FIFO.
%Note the memory ``process", which does not block on reads,
%can block on writes (responses to other processes), when a large amount of data returned are not consumed promptly. It may therefore also contribute to the development of artificial deadlocks. 
In a process network, % generated from a sequential program, 
these deadlocks, if left unresolved, would cause a mismatch between
the resultant behavior of the original code and the network. To tackle this problem, many approaches were proposed over the years, most of which involves monitoring execution and dynamically incrementing FIFO sizes when deadlock occurs~\cite{parks1995bounded}\cite{Geilen:2003:REK:1765712.1765736}\cite{4217338},
or only targets highly regular kernels.
%This is due to the infeasibility of analyzing deadlock for generic process networks at compile time. 
In our flow, we ensure, by construction, each FIFO is only read/written by a single instruction in each process, which makes the generated PN easier to analyze than a generic PN. We can therefore use a static analysis framework to find the needed FIFO sizes.

@@

\subsection{High Level Synthesis of Accelerators}
\label{pre2}
HLS for FPGAs are being active developed in both industry~\cite{tools:vivadohls} and academia~\cite{tools:legup}.
These HLS tools attempt to capture parallelism in a function written with high level languages like C/C++. With the help of various user provided guidance, it generates hardware in the form of RTL. 
%Compute operations and data accesses are allocated and scheduled according to the %dependency constraints extracted from the input program, and the resource %constraints imposed by the target platform. 
The synthesized circuits usually follow a Finite State Machine with Datapath (FSMD) paradigm. Activation of a particular operator in the datapath is associated
with a certain clock cycle and the execution of the computation engine is orchestrated by a synthesized central controller.

@@

The insertion of the sender/receiver of the barrier happens after the instruction partitioning. Depending on if a partition is the sender or receiver of the token,
an extra store/load operation is introduced into the set of instructions, associated
with the basic block where the barrier occurs.
\end{comment}
\subsubsection{Optimizing Communication v.s. Computation}
\label{opt}
 Regardless of exactly how the instructions are partitioned, as each process performs its computation locally, the necessity to communicate with its peers can incur implementation overheads.
For each communication channel, buffer space is to be allocated, and sender/receiver primitives also need to be inserted. Therefore, it is sometimes beneficial to duplicate some nodes after the instruction partitioning. Given sets of instruction nodes created after applying a platform-specific algorithm, for each of these set $G$, the problem of whether to copy other instruction nodes into $G$ can be formulated as an integer programming problem.

@@

Figure~\ref{fig:example} is an example PN we can use to illustrate this problem. The original loop is converted to a two-process network
with two single slot communication channels.
Figure~\ref{fig:2schedules} then shows the scheduling and communication of the processes. Each instruction invocation is prefixed with a vector, representing its place in the (j,i) iteration space, and three types of dependency edges are used
to show the necessary precedence among these invocations. There is no cycle of
dependency in (a), and therefore no deadlocks, when both processes execute in program order. However, in (b), 
a common HLS technique, loop pipelining, is applied to \textit{Process 1}, where
iterations of the loop are aggressively overlapped. Every cycle, a new iteration is started. Because of the long latency of load from \textit{in[i]} and
the multiplication, the sending of the first \textit{r2} token, $_{(0,0)}$store(r2Q,r2), is scheduled
after many instruction invocations for later iterations. Due to the static scheduling, it can not execute until after $_{(0,2)}$store(r1Q,r1), whose
writing to FIFO r1Q cannot happen until the previous token is consumed by 
$_{(0,1)}$load(r1Q). Because of the scheduling in process 2, this load from r1Q
occurs after $_{(0,0)}$load(r2Q), which cannot complete unless $_{(0,0)}$store(r2Q,r2)
can send the token.  An artificial deadlock is therefore created.

@@

\begin{figure}[htp]
\begin{center}
\includegraphics[width=0.8\linewidth]{fig/predGraphNew.pdf}
\caption{Precedence Graph for Process Network
\label{fig:predgraph}}
\end{center}
\vspace{-2.5em}
\end{figure}

@@

Resolving
the deadlock involves choosing one of the $DepF$ edges
in the cycle and add enough slots (making the weight more negative)
such that the weight of the cycle gets a negative leading element. For our example in figure~\ref{fig:predgraph}, if the capacity of r1Q
is increased by 1, the weight on its corresponding $DepF$ becomes (0,-2). Consequently, the identified cycle in the figure would have
a total weight (0,-1), when the original artificial deadlock disappears.
In general, for every deadlock causing cycle, the $DepF$ edge whose channel has the smallest width can be selected for capacity expansion, so the cost of these extra slots is minimized. This heuristic does not necessarily produce an absolute
global optimal solution in the presence of multiple deadlock causing cycles, but suffices for our use cases.

%\subsection{Modelling Burst Mode Memory Access}

\subsection{Deadlocks in Memory Accesses}
To exploit the memory bandwidth more efficiently, burst mode memory accesses are often used in accelerators. Even though the memory ``process" does not block on reads, if the response data it sends to other normal processes is not consumed promptly, it can blocked on writes. 
The head-of-line blocking can then create a dependency cycle, resulting in a deadlock. Our precedence graph can be used to compute necessary buffer size to resolve this situation as well.

Figure~\ref{fig:memBurst} illustrates how multiple burst mode memory accesses can
create a deadlock. The reaction of the memory is modelled by a pop action
responding to the incoming request, and a push action producing one response data token. Meanwhile, the sending of memory requests and receiving of responses are decoupled and assigned to different slots in the process schedule. This is often done in HLS because of the long latency of memory accesses.
For burst mode requests associated with a loop, the request sending is moved to outside of the loop as shown in the figure. The symbolic variable $B$ in the weight of the $DepM$ edge signifies how many tokens have to be consumed/buffered before the response for the next request ($data2$) become accessible for the response receiver. Assume the arbiter for the memory subsystem implements a policy which can potentially take in and serve multiple ($K$) consecutive requests from one stream, $B$ can be as big as $BurstSize_{max}*K$. 
In the case of a round-robin arbiter, $B = BurstSize_{max}$ and for our experiment platform which uses AXI bus protocol, this number is 256. With the precedence graph, we can easily compute that the buffering for the two memory response channels needs to be increased to $B$ for the deadlocks to be resolved.
\begin{figure}[htp]
\begin{center}
\includegraphics[width=1.05\linewidth]{fig/burstMem.pdf}
\caption{Deadlock in Burst Memory Accesses
\label{fig:memBurst}}
\end{center}
\vspace{-3.0em}
\end{figure}

@@

@@

The FPGA device used for our evaluation is Xilinx's Zynq-7000 XC7Z020 FPGA
SoC. It contains two parts: an ARM based processing
system (PS), and the programmable logic (PL). 
The baseline for our evaluation is the performance of each kernel executing
on the ARM core, which is an out-of-order, dual-issue hard processor running at 667MHz. Zynq also provides two
options for the accelerators in PL to access the main memory: 
through the accelerator coherence port (ACP), or the high performance (HP) port. 
ACP connects to the snoop control unit in the PS, and 
thus accesses the on-chip cache of the PS. On the other hand, the HP port connects
directly to the memory controller, thus often incurs longer latency
for data requests.

@@

\end{tabular}
\vspace{-2.0em}
\label{tab:datasize}
\end{table}

\begin{figure*}[htp]
\begin{center}
\includegraphics[width=0.95\linewidth]{fig/perfNew.pdf}
\caption{Performance of Conventional Accelerators and Accelerator Networks
\label{fig:acpraw}}
\end{center}
\vspace{-2.9em}
\end{figure*}

In our study, Vivado HLS, a state-of-the-art high level synthesis tool provided by Xilinx,
is used to create the accelerators for the original sequential code, as well as the processes generated by our flow. With the target clock period set to 
8ns during HLS, the tightest timing constraints post place \& route implementations managed to meet range from 111 to 150MHz. 
All design points shown in this section 
use the highest achievable frequency as the actual operating clock frequency. 
%\vspace{-1.0em}
\subsection{Performance Comparisons}
In figure~\ref{fig:acpraw}, 
performance of the different implementations are presented. 
Conventional accelerators and accelerator networks with different memory subsystem
configurations are compared. All the numbers are normalized to the baseline.
Design points with caches are using Xilinx System Cache IP, configured to be 64KB and 2-way associative.

@@

@@

%\subsection{Area and Energy Comparison}
%\vspace{-1.0em}
\subsection{Area comparison}
To quantify the impact of our proposed methodology on area, 
we have compared the FPGA resouce usage of conventional accelerators and
the accelerator networks. Table~\ref{tab:areacom} shows the results, 
where each implementation is complemented with two different memory subsystem configurations.
\renewcommand{\tabcolsep}{0.5pt}
\begin{table}[htbp]
%\vspace{-1.2em}
\caption{Resource Usage of Accelerators.}
%\vspace{-1.2em}
\scriptsize
\centering
\begin{tabular}{| c | c | c | c | c | c | c | c| }
  \hline            
  \multirow{2}{*}{} &  &   \multicolumn{3}{c|}{\bf ACP  } & \multicolumn{3}{c|}{\bf ACP + 64KB Cache }   \\
 \cline{3-8} 
 {\bf Benchmark}   &    & LUT& FFs& BRAM & LUT &   FFs      & BRAM    \\
  %
  %                 & LUT &FFs&  BRAM &  LUT &   FFs      & BRAM   \\
  \hline            
  \hline
